---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
 

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
head(activity)
summary(activity)
```
## What is mean total number of steps taken per day?
```{r}
dailysteps <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)

```
Calculating mean number of steps
```{r}
meanstepsdaily <- mean(dailysteps, na.rm = TRUE)
meanstepsdaily
```
Calculating median number of steps
```{r}
medstepsdaily <- median(dailysteps, na.rm = TRUE)
medstepsdaily
```

### Histogram for total number of steps taken each day ###
```{r}
hist(dailysteps, breaks = 10, xlab = "daily step count", main = "Histogram of daily steps", col = "darkorchid")
```

## What is the average daily activity pattern?
Calculate number of steps each day
```{r}
activity$date <- as.character(activity$date)
meanstepsinterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```

### Time-series plot of mean steps per 5-minute interval ###
```{r}
plot(row.names(meanstepsinterval), meanstepsinterval, type = "l", xlab = "5 minute intervals (12AM - 11:59PM)", ylab="mean number of steps", main="Mean steps per interval", col="darkorchid", cex=1, cex.axis=0.75, cex.lab=0.75, cex.main=0.95, font.lab=2, font=2, lwd = 2)
```
Which interval averages the most steps?
```{r}
maxstepsinterval <- which.max(meanstepsinterval)
print(maxstepsinterval)
names(maxstepsinterval)
```
## Imputing missing values
```{r}
sum(is.na(activity))
```
*Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.*

The mean number of steps for each interval is used to impute the missing values (NAs) in the data.
```{r}
imputed <- activity
meansteps <- mean(activity$steps, na.rm=TRUE)
imputed$steps[is.na(imputed$steps)] <- meansteps
```
Are there any NAs remaining in the dataset?
```{r}
sum(is.na(imputed))
```
New table with NAs imputed from interval means
```{r}
dailystepsimputed <- tapply(imputed$steps, imputed$date, sum)
head(dailystepsimputed)
summary(imputed)
```
### Histogram for total number of steps taken each day with NA values imputed from interval means ###
```{r}
hist(dailystepsimputed, breaks = 10, xlab = "daily step count", main = "Histogram of daily steps (imputed means)", col = "orchid2")
```
Calculate new mean number of steps
```{r}
imputedmeanstepsdaily <- mean(dailystepsimputed)
imputedmeanstepsdaily
```
Calculate new median number of steps
```{r}
imputedmedstepsdaily <- median(dailystepsimputed)
imputedmedstepsdaily
```
Change in steps with imputed NA values?
```{r}
summary(dailysteps)
```

```{r}
summary(dailystepsimputed)
```
How did the imputed data differ from the original data?
```{r}
meandifference <- mean(dailystepsimputed) - mean(dailysteps)
meandifference
```

```{r}
mediandifference <- median(dailystepsimputed) - median(dailysteps)
mediandifference
```
## Are there differences in activity patterns between weekdays and weekends?

```{r}
imputed$daysweek <- weekdays(as.Date(imputed$date))
imputed$days[(imputed$daysweek == "Saturday" | imputed$daysweek == "Sunday")] <- "weekend"
imputed$days[!(imputed$daysweek == "Saturday" | imputed$daysweek == "Sunday")] <- "weekday"
head(imputed)
imputed$daysweek <- as.factor(imputed$daysweek)
imputed$days <- as.factor(imputed$days)
```

### Time-series panel plot for mean number of steps taken between weekdays & weekends ###
```{r}
library(lattice)
imputedweekdaymeans <- aggregate(steps ~ interval + days, data = imputed, mean)
xyplot(steps ~ interval | days, data = imputedweekdaymeans, type = "l", xlab = "interval", 
       ylab = "mean number of steps", main = "Mean number of step per 5 minute interval", col = "orchid", lwd = 2, layout = c(1,2))
```

